Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-01-20 manuscript/manuscript.Rmd
Update manuscript.Rmd

For modified transformations show:
- Comparison with figure 1.
- A shifted left-skewed distribution, with the same shifts.
- A shifted right-skewed distribution, with the same shifts.

```{r shifted-distributions, echo=FALSE, fig.cap=cap}
cap <- paste0(
  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "Unlike the orginal power transformations, $\\lambda = 1$ is consistently found using shift-sensitive power transformations.")

@@

  p <- ggplot2::ggplot(
    data = data,
    mapping = ggplot2::aes(
      x = mu,
      y = lambda,
      colour = method,
      shape = version))
  p <- p + plot_theme
  p <- p + ggplot2::geom_point()
  p <- p + ggplot2::geom_hline(
    yintercept = 1.0, 
    linetype="longdash", 
    colour="grey40")
  p <- p + ggplot2::scale_x_continuous(
    name = latex2exp::TeX("$\\mu$"),
    labels = scales::math_format())
  p <- p + ggplot2::scale_y_continuous(
    name = latex2exp::TeX("$\\lambda$"),
    limits = limits)

@@

data <- .create_lambda_shift_data(
  x = x,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_normal <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(-3.6, 1.1))
p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_normal <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(-3.6, 1.1))
p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_yj_normal <- p_yj_normal + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

@@

data <- .create_lambda_shift_data(
  x = x_right_skewed,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_right <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(1.0, 65))
p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_right <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(1.0, 65))
p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE)
p_yj_right <- p_yj_right + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

@@

data <- .create_lambda_shift_data(
  x = x_left_skewed,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_left <- .create_lambda_shift_plot(
  data=data[method == "Box-Cox"],
  plot_theme=plot_theme,
  limits = c(-10.5, 0.5))
p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
  palette="ggthemes::Tableau_10",
  drop=FALSE,
  guide="none")
p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_left <- .create_lambda_shift_plot(
  data=data[method == "Yeo-Johnson"],
  plot_theme=plot_theme,
  limits = c(-10.5, 0.5))
p_yj_left <- p_yj_left + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_yj_left <- p_yj_left + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

@@

>>>>

To assess whether the modified power transformations lead to $\lambda$ values that are invariant to shifts, we simulate three different distributions. The first distribution is a normal distribution, as previously shown in Figure **INSERT REFERENCE**. We first randomly draw $10000$ values from a normal distribution: $x_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$. The second distribution is a right-skewed distribution $x_{\text{right}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) + 1\right)^{\frac{1}{10}}$. The third distribution is a left-skewed distribution $x_{\text{left}} = \left( x_{\text{normal}} - \text{min} \left(x_{\text{normal}}\right) \right)^3$. Subsequently, all three distributions are standardised to have mean $0$ and standard deviation $1$. We then compute transformation parameter $\lambda$ using the original definitions (Eq. **INSERT REFERENCE** and **INSERT REFERENCE**) and the modified definitions (Eq. **INSERT REFERENCE** and **INSERT REFERENCE**) for each distribution, after adding a positive value $d$ to each distribution. Here, $d \in [1, 10^6]$.

```{r shifted-distributions, echo=FALSE, fig.cap=cap}
cap <- paste0(
  "Power transformations modified to be shift-sensitive produce transformation parameters that are invariant to location.")

@@

  p <- ggplot2::ggplot(
    data = data,
    mapping = ggplot2::aes(
      x = mu,
      y = lambda,
      colour = method,
      shape = version))
  p <- p + plot_theme
  p <- p + ggplot2::geom_point()
  p <- p + ggplot2::geom_hline(
    yintercept = 1.0, 
    linetype="longdash", 
    colour="grey40")
  p <- p + ggplot2::scale_x_continuous(
    name = latex2exp::TeX("$d$"),
    labels = scales::math_format())
  p <- p + ggplot2::scale_y_continuous(
    name = latex2exp::TeX("$\\lambda$"),
    limits = limits)

@@

data_normal <- .create_lambda_shift_data(
  x = x,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_normal <- .create_lambda_shift_plot(
  data = data_normal[method == "Box-Cox"],
  plot_theme = plot_theme,
  limits = c(-3.6, 1.1))
p_bc_normal <- p_bc_normal + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_bc_normal <- p_bc_normal + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_normal <- .create_lambda_shift_plot(
  data=data_normal[method == "Yeo-Johnson"],
  plot_theme = plot_theme,
  limits = c(-3.6, 1.1))
p_yj_normal <- p_yj_normal + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_yj_normal <- p_yj_normal + ggplot2::theme(
  axis.title.y = ggplot2::element_blank(),
  axis.text.y = ggplot2::element_blank())

@@

data_right <- .create_lambda_shift_data(
  x = x_right_skewed,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_right <- .create_lambda_shift_plot(
  data = data_right[method == "Box-Cox"],
  plot_theme = plot_theme,
  limits = c(1.0, 65))
p_bc_right <- p_bc_right + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_bc_right <- p_bc_right + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_right <- .create_lambda_shift_plot(
  data = data_right[method == "Yeo-Johnson"],
  plot_theme = plot_theme,
  limits = c(1.0, 65))
p_yj_right <- p_yj_right + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE)
p_yj_right <- p_yj_right + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

@@

data_left <- .create_lambda_shift_data(
  x = x_left_skewed,
  shift_range = shift_range)

@@

# Box-Cox
p_bc_left <- .create_lambda_shift_plot(
  data = data_left[method == "Box-Cox"],
  plot_theme = plot_theme,
  limits = c(-10.5, 0.5))
p_bc_left <- p_bc_left + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_bc_left <- p_bc_left + ggplot2::scale_shape_discrete(guide="none")

# Yeo-Johnson
p_yj_left <- .create_lambda_shift_plot(
  data = data_left[method == "Yeo-Johnson"],
  plot_theme = plot_theme,
  limits = c(-10.5, 0.5))
p_yj_left <- p_yj_left + paletteer::scale_color_paletteer_d(
  palette = "ggthemes::Tableau_10",
  drop = FALSE,
  guide = "none")
p_yj_left <- p_yj_left + ggplot2::theme(
  axis.title.y=ggplot2::element_blank(),
  axis.text.y=ggplot2::element_blank())

#### Aggregate -----------------------------------------------------------------

data_normal[, "distribution":="normal"]
data_left[, "distribution":="left-skewed"]
data_right[, "distribution":="right-skewed"]

aggregate_data <- data.table::rbindlist(list(data_normal, data_left, data_right))[, list("mean_lambda"=mean(lambda), "sd_lambda"=stats::sd(lambda)), by=c("method", "version", "distribution")]

@@

The result is shown in Figure **INSERT REFERENCE**. For each distribution, transformation parameter $\lambda$ varies with $d$ when estimated using the original definitions. In contrast, estimations of $\lambda$ modified power transformations were invariant to $d$.
